Size class small comparisons

### 2020-08-05 analysis/reports/manuscript_main.Rmd
some updates to manusciprt and manuscript figures

ggplot(filter(all_di, s0 > 2), aes(skew_percentile)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 95, color = "red") +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") + ggtitle("Skewness, all sites")

ggplot(filter(all_di, s0 > 2, log_nparts > log(50)), aes(skew_percentile)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 95, color = "red") +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") + ggtitle("Skewness, all sites")

ggplot(all_di, aes(simpson_percentile)) + 
 geom_histogram(bins = 100) +
  geom_vline(xintercept = 5, color = "red") +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") + ggtitle("Evenness, all sites")
ggplot(filter(all_di, log_nparts > log(50)), aes(simpson_percentile)) + 
 geom_histogram(bins = 100) +
  geom_vline(xintercept = 5, color = "red") +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") + ggtitle("Evenness, all sites")
```

@@

overall_skew_results <- all_di %>%
  filter(s0 > 2) %>%
  mutate(high_skew = skew_percentile > 95) %>%
  group_by(dat) %>%
  summarize(proportion_skew_high = mean(high_skew),
            nsites_skew = dplyr::n())

@@

overall_skew_results2 <- all_di %>%
  filter(s0 > 2, log_nparts > log(50)) %>%
  mutate(high_skew = skew_percentile > 95) %>%
  group_by(dat) %>%
  summarize(proportion_skew_high = mean(high_skew),
            nsites_skew = dplyr::n())

overall_simpson_results2 <- all_di %>%
  filter(log_nparts > log(50)) %>%
  mutate(low_even = simpson_percentile < 5) %>%
  group_by(dat) %>%
  summarize(proportion_even_low = mean(low_even),
            nsites_even = dplyr::n())

@@

There are fewer sites for skewness because we exclude sites with only 2 species.

# Size of feasible set for each dataset

```{r fs size hists}

all_di <- all_di %>%
  mutate(log_nparts_truncated = ifelse(
    log_nparts > 50,
    50,
    log_nparts)
  )

ggplot(all_di, aes(log_nparts_truncated)) +
  geom_histogram() +
  facet_wrap(vars(dat), scales = "free_y") +
  theme_bw()

fs_size_results <- all_di %>%
  group_by(dat) %>%
  summarize(prop_smaller_than_5=  mean(log_nparts <= 5),
            prop_smaller_than_7 =  mean(log_nparts <= 7),
            prop_smaller_than_10 = mean(log_nparts <= 10))
fs_size_results

@@

FIA is the only dataset that is overwhelmingly dominated by sites with fewer than `r exp(10)` elements in the feasible set. MCDB also has many small sites, but these are counterbalanced by larger ones.

62% of the FIA sites have fewer than `r exp(7)` elements in the feasible set, and 30% have fewer than `r exp(5)`.

```{r fs size to s0 and n0, fig.dim = c(3,3)}
ggplot(all_di, aes(x = log_s0, y = log_n0, color = log_nparts <= 5)) +
  geom_point() +
  theme_bw() +
  theme(legend.position = "top")

ggplot(all_di, aes(x = log_s0, y = log_n0, color = log_nparts <= 7)) +
  geom_point() +
  theme_bw()+
  theme(legend.position = "top")

ggplot(all_di, aes(x = log_s0, y = log_n0, color = log_nparts <= 10)) +
  geom_point() +
  theme_bw()+
  theme(legend.position = "top")
```

Having fewer than 150 elements ~ fewer than 3 species, or fewer than `r exp(2.75)` individuals  with up to 20 species.

Fewer than 1000 elements ~ fewer than 3 species, or fewer than `r exp(3.25)` individuals with up to 20 species.

Fewer than 20000 elements ~ fewer than 4 species, or fewer than `r exp(4)` individuals.

# 95% interval for small versus large communities

```{r 95 interval}

ggplot(all_di, aes(x = log_nparts, y = skew_95_ratio_1t)) +
  geom_point(alpha = .2) + theme_bw() + ggtitle("Skewness 95% interval")

ggplot(all_di, aes(x = log_nparts, y = skew_95_ratio_1t)) +
  geom_point(alpha = .01) +
  xlim(0, 15) +
  theme_bw() +
  ggtitle("Skewness 95% interval - only small feasible sets")

ggplot(all_di, aes(x = log_nparts, y = simpson_95_ratio_1t)) +
  geom_point(alpha = .2) + theme_bw()+ ggtitle("Evenness 95% interval")

ggplot(all_di, aes(x = log_nparts, y = simpson_95_ratio_1t)) +
  geom_point(alpha = .05) +
  xlim(0, 15) +
  theme_bw()+
  ggtitle("Evenness 95% interval - only small feasible sets")

The 95% ratio decreases for large FS.

For skewness, it is still quite high even up to 20,000 elements in the FS. But it is starting to decline by then?

# Comparison of FIA to similarly-sized communities

```{r fia sized}
fia_max_s <- max(filter(all_di, dat == "fia")$s0)
fia_max_n <- max(filter(all_di, dat == "fia")$n0)

fia_max_nparts <-max(filter(all_di, dat == "fia")$log_nparts)

fia_sized <- filter(all_di, s0 <= fia_max_s, n0 <= fia_max_n) %>%
  mutate(fia_yn = dat == "fia") %>%
  mutate(fia_yn = ifelse(fia_yn, "FIA", "Other datasets")) %>%
  mutate(right_corner = FALSE)

ggplot(filter(fia_sized, s0 > 2, nparts > 50), aes(skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y")

ggplot(fia_sized, aes(x = log_nparts)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  geom_vline(xintercept = log(50))

ggplot(fia_sized, aes(s0, n0, color = fia_yn)) +
  geom_point()

# Removing the smallest communities from FIA

```{r fia no small}

ggplot(filter(fia_sized, s0 > 2, nparts > 50, nparts < exp(7)), aes(skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("50-1,000 parts in FS")

ggplot(filter(fia_sized, s0 > 2, nparts > 50, nparts >= exp(7)), aes(skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("More than 1,000 parts in FS")

ggplot(filter(fia_sized, s0 > 2, nparts > 50, nparts >= exp(10)), aes(skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  ggtitle("More than 20,000 parts in FS")
```

>>>>

ggplot(filter(all_di, s0 > 2), aes(skew_percentile_excl)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 95, color = "red") +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") + ggtitle("Skewness, all sites")

ggplot(all_di, aes(simpson_percentile)) + 
 geom_histogram(bins = 100) +
  geom_vline(xintercept = 5, color = "red") +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y") + ggtitle("Evenness, all sites")

@@

overall_skew_results <- all_di %>%
  filter(s0 > 2) %>%
  mutate(high_skew = skew_percentile_excl > 95) %>%
  group_by(dat) %>%
  summarize(proportion_skew_high = mean(high_skew),
            nsites_skew = dplyr::n())

@@

overall_skew_results2 <- all_di %>%
  mutate(fia_yn = ifelse(dat == "fia", "fia", "not fia")) %>%
  filter(s0 > 2) %>%
  mutate(high_skew = skew_percentile_excl > 95) %>%
  group_by(fia_yn) %>%
  summarize(proportion_skew_high = mean(high_skew),
            nsites_skew = dplyr::n())

overall_simpson_results2 <- all_di %>%
   mutate(fia_yn = ifelse(dat == "fia", "fia", "not fia")) %>%
  mutate(low_even = simpson_percentile < 5) %>%
  group_by(fia_yn) %>%
  summarize(proportion_even_low = mean(low_even),
            nsites_even = dplyr::n())

@@

```{r 95 intervals}

ggplot(all_di, aes(x = log_nparts, y= skew_95_ratio_1t)) +
  geom_point(alpha = .1) +
  theme_bw() +
  xlim(0, 100)

ggplot(all_di, aes(x = log_nparts, y= simpson_95_ratio_1t)) +
  geom_point(alpha = .1) +
  theme_bw() +
  xlim(0, 100)

@@

### 2020-08-05 analysis/reports/smallness.Rmd
some updates to manusciprt and manuscript figures

@@

@@

small_di %>%
  filter(nparts< maxparts, nparts > minparts,
         s0 > 2) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high = mean(skew_percentile_excl > 95),
            nskew_sites = dplyr::n())

@@

small_di %>%
  filter(nparts< maxparts, nparts > minparts,
         s0 > 2) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high = mean(skew_percentile_excl > 95),
            nskew_sites = dplyr::n())

@@

small_di %>%
  filter(nparts< maxparts, nparts > minparts,
         s0 > 2) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high = mean(skew_percentile_excl > 95),
            nskew_sites = dplyr::n())

>>>>

this_di <- small_di %>%
            filter(nparts > minparts,
                   nparts < maxparts)

        ggplot(small_di, aes(x = (s0), y = (n0))) +
            geom_point() +
            geom_point(data = this_di, color = "blue") +
            theme_bw() +
            facet_wrap(vars(fia_yn))

@@

 ggplot(small_di, aes(log_nparts)) +
            geom_histogram(bins = 30) +
            theme_bw() +
            facet_wrap(vars(fia_yn), scales = "free_y") +
            geom_vline(xintercept = c(log(minparts),
                                      log(maxparts)), color = "red")

@@

ggplot(filter(small_di,
              nparts< maxparts, nparts > minparts), aes(skew_95_ratio_1t)) +
         geom_histogram() +
         theme_bw()

ggplot(filter(small_di,
              nparts< maxparts, nparts > minparts), aes(simpson_95_ratio_1t)) +
         geom_histogram() +
         theme_bw()
small_di %>%
  filter(nparts< maxparts, nparts > minparts,
         s0 > 2) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high = mean(skew_percentile_excl > 95),
            nskew_sites = dplyr::n())

@@

this_di <- small_di %>%
            filter(nparts > minparts,
                   nparts < maxparts)

        ggplot(small_di, aes(x = (s0), y = (n0))) +
            geom_point() +
            geom_point(data = this_di, color = "blue") +
            theme_bw() +
            facet_wrap(vars(fia_yn))

ggplot(filter(small_di,
              nparts< maxparts, nparts > minparts,
              s0 > 2), aes(skew_percentile_excl)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  theme_bw()
 ggplot(small_di, aes(log_nparts)) +
            geom_histogram(bins = 30) +
            theme_bw() +
            facet_wrap(vars(fia_yn), scales = "free_y") +
            geom_vline(xintercept = c(log(minparts),
                                      log(maxparts)), color = "red")
ggplot(filter(small_di,
              nparts< maxparts, nparts > minparts), aes(simpson_percentile)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  theme_bw()

ggplot(filter(small_di,
              nparts< maxparts, nparts > minparts), aes(skew_95_ratio_1t)) +
         geom_histogram() +
         theme_bw()

ggplot(filter(small_di,
              nparts< maxparts, nparts > minparts), aes(simpson_95_ratio_1t)) +
         geom_histogram() +
         theme_bw()
small_di %>%
  filter(nparts< maxparts, nparts > minparts,
         s0 > 2) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high = mean(skew_percentile_excl > 95),
            nskew_sites = dplyr::n())

@@

 ggplot(small_di, aes(log_nparts)) +
            geom_histogram(bins = 30) +
            theme_bw() +
            facet_wrap(vars(fia_yn), scales = "free_y") +
            geom_vline(xintercept = c(log(minparts),
                                      log(maxparts)), color = "red")
ggplot(filter(small_di,
              nparts< maxparts, nparts > minparts), aes(simpson_percentile)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(fia_yn), scales = "free_y") +
  theme_bw()

ggplot(filter(small_di,
              nparts< maxparts, nparts > minparts), aes(skew_95_ratio_1t)) +
         geom_histogram() +
         theme_bw()

ggplot(filter(small_di,
              nparts< maxparts, nparts > minparts), aes(simpson_95_ratio_1t)) +
         geom_histogram() +
         theme_bw()
small_di %>%
  filter(nparts< maxparts, nparts > minparts,
         s0 > 2) %>%
  group_by(fia_yn) %>%
  summarize(prop_skew_high = mean(skew_percentile_excl > 95),
            nskew_sites = dplyr::n())
